DEG_RNA_summary

### 2025-02-25 05b_Functional_enrichment_analysis/DEG_gProfiler.Rmd
cleaned up code

```{r setup}
library(gprofiler2)
library(plyr)
library(dplyr)
library(ggplot2)
library(tibble)
library(cowplot)

@@

length(RNAlistRDSdf$gene)
RNAlistRDSdf_filt <- RNAlistRDSdf %>% dplyr::filter(AveExpr > 18.53202) 
length(RNAlistRDSdf_filt$gene)
```

@@

gProfiler based on RNA/SCT assay and per annotation with gene expression >18.53202
```{r}
##function to run gProfiler and save output
gProfiler_FUN <- function(x,ANALYSIS,GRADE,ORDERED_QUERY,ASSAY,dir_out) {
#gProfiler
custom_bg <- unique(x$gene)
query <- x %>%
  dplyr::filter(AveExpr > 18.53202) %>%
  dplyr::filter(p_adj.loc < 0.05) %>%
  dplyr::arrange(p_adj.loc) #arrange by lowest p_adj.loc
if(nrow(query) > 0){
query <- distinct(query, gene, .keep_all= TRUE) #remove duplicates of gene with higher p_adj.loc
query <- query$gene
gostres <<- gost(query = query, 
                organism = "hsapiens", 
                ordered_query = ORDERED_QUERY, #can use TRUE, p_adj.loc has been arranged 
                multi_query = FALSE, 
                significant = TRUE, #only statistically significant results should be returned
                exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, 
                evcodes = TRUE,
                user_threshold = 0.05, 
                correction_method = "g_SCS", #g_SCS is default
                domain_scope = "annotated", #if custom_bg is set, I believe this shouldn't matter too much
                custom_bg = custom_bg, #custom background, use!
                numeric_ns = "", 
                sources = NULL, #by default all sources are analyzed
                as_short_link = FALSE)
if(length(gostres) > 0){
results <- gostres$result
results$cluster_id <- levels(as.factor(x$cluster_id))
results$analysis <- print(ANALYSIS)
results$assay <- print(ASSAY)
results$grade <- print(GRADE) 
results$ordered_query <- print(ORDERED_QUERY)
results$parents <- NULL
write.csv(results,paste0("gProfiler_",ANALYSIS,"_",GRADE,"_",levels(as.factor(x$cluster_id)),"_OrderedQuery",as.character(ORDERED_QUERY),".csv"))
print(paste0("Saved gotres as gProfiler_",ANALYSIS,"_",GRADE,"_",levels(as.factor(x$cluster_id)),"_OrderedQuery",as.character(ORDERED_QUERY),".csv"))
plot <- gostplot(gostres, capped = TRUE, interactive = FALSE) + ggtitle(paste0("gProfiler_",ANALYSIS,"_",GRADE,"_",levels(as.factor(x$cluster_id)),"_OrderedQuery",as.character(ORDERED_QUERY)))
ggsave(plot = plot, filename = paste0(dir_out,"gProfiler_",ANALYSIS,"_",GRADE,"_",levels(as.factor(x$cluster_id)),"_OrderedQuery",as.character(ORDERED_QUERY),".png"), width = 7, height = 5)
print("Saved plot")
}
}
}
##function to run gProfiler_FUN on different grades, analysis and assay
RUN_gProfier <- function(GRADE, ANALYSIS,ASSAY,SPECIFICATION){
dir_out <- paste0("/home/common/data/output/projects/ENDO/E044/A031/gProfiler/E044",ANALYSIS,"/",ASSAY,"assay_",SPECIFICATION,"/")
dir.create(dir_out, recursive = TRUE)
setwd(dir_out)
#Read paths to DEG output .rds files
if (ASSAY == "RNA") { 
    PATHS <- list.files(path = paste0("/home/common/data/output/projects/ENDO/E044/",ANALYSIS), pattern = "res_mm_dream.rds", recursive = TRUE, full.names = TRUE)
    } else if (ASSAY == "SCT") {
    PATHS <- list.files(path = paste0("/home/common/data/output/projects/ENDO/E044/",ANALYSIS), pattern = "mm_dream_sct_counts_all.rds", recursive = TRUE, full.names = TRUE)
    }
#Select paths based on all/mild/severe
if (GRADE == "all") { 
    PATHS <- PATHS[!grepl("_severe", PATHS)]
    PATHS <- PATHS[!grepl("_mild", PATHS)]
    print(PATHS)
    } else if (GRADE == "mild") {
    PATHS <- PATHS[grepl("_mild", PATHS)]
    print(PATHS)
    } else if  (GRADE == "severe") {
    PATHS <- PATHS[grepl("_severe", PATHS)]
    print(PATHS)
    } 
RNAlistRDS <- lapply(PATHS, readRDS)
RNAlistRDS <- unlist(RNAlistRDS, recursive = FALSE) #combine sub-lists
RNAlistRDS <- RNAlistRDS[!duplicated(RNAlistRDS)]
lapply(RNAlistRDS, gProfiler_FUN, ANALYSIS = ANALYSIS,GRADE = GRADE,ORDERED_QUERY = FALSE, ASSAY = ASSAY, dir_out = dir_out)
lapply(RNAlistRDS, gProfiler_FUN, ANALYSIS = ANALYSIS,GRADE = GRADE,ORDERED_QUERY = TRUE, ASSAY = ASSAY, dir_out = dir_out)
}

RUN_gProfier(GRADE = "all", ANALYSIS = "A030", ASSAY = "RNA",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
RUN_gProfier(GRADE = "severe", ANALYSIS = "A030", ASSAY = "RNA",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
RUN_gProfier(GRADE = "mild", ANALYSIS = "A030", ASSAY = "RNA",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
RUN_gProfier(GRADE = "all", ANALYSIS = "A030", ASSAY = "SCT",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
RUN_gProfier(GRADE = "severe", ANALYSIS = "A030", ASSAY = "SCT",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
RUN_gProfier(GRADE = "mild", ANALYSIS = "A030", ASSAY = "SCT",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
```

##Ongoing xxxx
gProfiler based on RNA assay and per annotation with gene expression >18.53202 and BH pval adjusted
```{r}
##function to run gProfiler and save output
gProfiler_FUN <- function(x,ANALYSIS,GRADE,ORDERED_QUERY,ASSAY,dir_out) {
#gProfiler
custom_bg <- unique(x$gene)
query <- x %>%
  dplyr::filter(AveExpr > 18.53202) %>%
  dplyr::filter(p_adj.loc < 0.05) %>%
  dplyr::arrange(p_adj.loc) #arrange by lowest p_adj.loc
if(nrow(query) > 0){
query <- distinct(query, gene, .keep_all= TRUE) #remove duplicates of gene with higher p_adj.loc
query <- query$gene
gostres <<- gost(query = query, 
                organism = "hsapiens", 
                ordered_query = ORDERED_QUERY, #can use TRUE, p_adj.loc has been arranged 
                multi_query = FALSE, 
                significant = TRUE, #only statistically significant results should be returned
                exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, 
                evcodes = TRUE,
                user_threshold = 0.05, 
                correction_method = "g_SCS", #g_SCS is default
                domain_scope = "annotated", #if custom_bg is set, I believe this shouldn't matter too much
                custom_bg = custom_bg, #custom background, use!
                numeric_ns = "", 
                sources = NULL, #by default all sources are analyzed
                as_short_link = FALSE)
if(length(gostres) > 0){
results <- gostres$result
results$cluster_id <- levels(as.factor(x$cluster_id))
results$analysis <- print(ANALYSIS)
results$assay <- print(ASSAY)
results$grade <- print(GRADE) 
results$ordered_query <- print(ORDERED_QUERY)
results$parents <- NULL
write.csv(results,paste0("gProfiler_",ANALYSIS,"_",GRADE,"_",levels(as.factor(x$cluster_id)),"_OrderedQuery",as.character(ORDERED_QUERY),".csv"))
print(paste0("Saved gotres as gProfiler_",ANALYSIS,"_",GRADE,"_",levels(as.factor(x$cluster_id)),"_OrderedQuery",as.character(ORDERED_QUERY),".csv"))
plot <- gostplot(gostres, capped = TRUE, interactive = FALSE) + ggtitle(paste0("gProfiler_",ANALYSIS,"_",GRADE,"_",levels(as.factor(x$cluster_id)),"_OrderedQuery",as.character(ORDERED_QUERY)))
ggsave(plot = plot, filename = paste0(dir_out,"gProfiler_",ANALYSIS,"_",GRADE,"_",levels(as.factor(x$cluster_id)),"_OrderedQuery",as.character(ORDERED_QUERY),".png"), width = 7, height = 5)
print("Saved plot")
}
}
}
##function to run gProfiler_FUN on different grades, analysis and assay
RUN_gProfier <- function(GRADE, ANALYSIS,ASSAY,SPECIFICATION){
dir_out <- paste0("/home/common/data/output/projects/ENDO/E044/A031/gProfiler/E044",ANALYSIS,"/",ASSAY,"assay_",SPECIFICATION,"/")
dir.create(dir_out, recursive = TRUE)
setwd(dir_out)
#Read paths to DEG output .rds files
if (ASSAY == "RNA") { 
    PATHS <- list.files(path = paste0("/home/common/data/output/projects/ENDO/E044/",ANALYSIS), pattern = "res_mm_dream.rds", recursive = TRUE, full.names = TRUE)
    } else if (ASSAY == "SCT") {
    PATHS <- list.files(path = paste0("/home/common/data/output/projects/ENDO/E044/",ANALYSIS), pattern = "mm_dream_sct_counts_all.rds", recursive = TRUE, full.names = TRUE)
    }
#Select paths based on all/mild/severe
if (GRADE == "all") { 
    PATHS <- PATHS[!grepl("_severe", PATHS)]
    PATHS <- PATHS[!grepl("_mild", PATHS)]
    print(PATHS)
    } else if (GRADE == "mild") {
    PATHS <- PATHS[grepl("_mild", PATHS)]
    print(PATHS)
    } else if  (GRADE == "severe") {
    PATHS <- PATHS[grepl("_severe", PATHS)]
    print(PATHS)
    } 
RNAlistRDS <- lapply(PATHS, readRDS)
RNAlistRDS <- unlist(RNAlistRDS, recursive = FALSE) #combine sub-lists
RNAlistRDS <- RNAlistRDS[!duplicated(RNAlistRDS)]
lapply(RNAlistRDS, gProfiler_FUN, ANALYSIS = ANALYSIS,GRADE = GRADE,ORDERED_QUERY = FALSE, ASSAY = ASSAY, dir_out = dir_out)
lapply(RNAlistRDS, gProfiler_FUN, ANALYSIS = ANALYSIS,GRADE = GRADE,ORDERED_QUERY = TRUE, ASSAY = ASSAY, dir_out = dir_out)
}

RUN_gProfier(GRADE = "all", ANALYSIS = "A030", ASSAY = "RNA",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
RUN_gProfier(GRADE = "severe", ANALYSIS = "A030", ASSAY = "RNA",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
RUN_gProfier(GRADE = "mild", ANALYSIS = "A030", ASSAY = "RNA",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
RUN_gProfier(GRADE = "all", ANALYSIS = "A030", ASSAY = "SCT",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
RUN_gProfier(GRADE = "severe", ANALYSIS = "A030", ASSAY = "SCT",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
RUN_gProfier(GRADE = "mild", ANALYSIS = "A030", ASSAY = "SCT",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
```

```{r Combine output tables from gProfiler}
#functions
summarize_gProfiler_PerAnnotation <- function(x){
ASSAY = "A030"
##Combine all results
gProfilerlist <- list.files(path = paste0("/home/common/data/output/projects/ENDO/E044/A031/gProfiler/E044",ASSAY,"/",x), pattern = "TRUE.csv$|FALSE.csv$", full.names = TRUE)
gProfilerlist <- lapply(gProfilerlist, read.csv) #read .rds
#Combine all gProfiler results into one df, save
gProfiler <- plyr::rbind.fill(gProfilerlist) #bind deg lists into one df
gProfiler$specification <- print(x)
#save csv
write.csv(gProfiler, paste0("/home/common/data/output/projects/ENDO/E044/A031/gProfiler/E044",ASSAY,"/gProfiler_",ASSAY,"_",x,"_summary.csv"))
table(gProfiler$grade)
#table(gProfiler$cluster_id,gProfiler$grade)
}

ASSAY = "A030"
DIRSLIST <- list.dirs(path = paste0("/home/common/data/output/projects/ENDO/E044/A031/gProfiler/E044",ASSAY),full.names = FALSE)
DIRSLIST_Merged <- DIRSLIST[grepl("Merged", DIRSLIST)]
DIRSLIST_PerAnnotation <- DIRSLIST[grepl("PerAnnotation", DIRSLIST)]
lapply(DIRSLIST_PerAnnotation,summarize_gProfiler_PerAnnotation)
```

@@

###interesting term_names
#Immune System and inflammation
ImmuneTermsGREP <- c("TGF", "TNF", "NOD", "IL-", "immun", "cytokine", "chemokine", "NF-kappa")
ImmuneTerms <- gProfiler_df %>%
  filter(grepl(paste(ImmuneTermsGREP, collapse = "|"), term_name, ignore.case = TRUE)) %>% 
  pull(term_name) %>% 
  unique()
#angiogenesis
angiogenesisTerms  <- c(unique(gProfiler_df$term_name[grepl("VEGF", gProfiler_df$term_name, ignore.case = TRUE)]),
                       unique(gProfiler_df$term_name[grepl("angiog", gProfiler_df$term_name, ignore.case = TRUE)]))
#extracellullar terms
extracellularTerms <- unique(gProfiler_df$term_name[grepl("extracellular", gProfiler_df$term_name, ignore.case = TRUE)])
#cell adhesion
adhesionTermsGREP <- c("adhesion", "junction")
adhesionTerms <- gProfiler_df %>%
  filter(grepl(paste(adhesionTermsGREP, collapse = "|"), term_name, ignore.case = TRUE)) %>% 
  pull(term_name) %>% 
  unique()
#growth factors
insulinTerms <- unique(gProfiler_df$term_name[grepl("insulin", gProfiler_df$term_name, ignore.case = TRUE)])
#Other interesting terms
otherTerms <- c('CTNNB1-DDX5-HDGF complex') #Hepatoma-Derived Growth Factor and DDX5 Promote Carcinogenesis and Progression of Endometrial Cancer by Activating β-Catenin
# Define the list of interesting terms
interesting_terms <- list(
  ImmuneTerms = ImmuneTerms,
  AngiogenesisTerms = angiogenesisTerms,
  ExtracellularTerms = extracellularTerms,
  AdhesionTerms = adhesionTerms,
  InsulinTerms = insulinTerms
  #,OtherTerms = otherTerms
)
# Loop through the list and assign categories
gProfiler_df$DEG_category <- NA
for (category_name in names(interesting_terms)) {
  terms <- interesting_terms[[category_name]]
  for (term in terms) {
    gProfiler_df$DEG_category[gProfiler_df$term_name == term] <- category_name
  }
}
table(gProfiler_df$DEG_category)

@@

##############################################################################################################################
##Merge the DEG table with the Terms

#RNA assay DEGs from proliferative samples
RNAlistRDS005_df <- read.csv("/home/common/data/output/projects/ENDO/E044/A031/RNAlistRDS005_unmerged.csv") %>%
  filter(annotation_reference == "Tan") %>%
  filter(AveExpr > 18.53202)
##Exclude Haemoglobin genes, mitochondrial genes, #ribosomal genes
hb.genes <- c("HBA1","HBA2","HBB","HBG2","HBG1","HBD","HBE1","HBZ")#haemoglobin genes, there are no hb genes in the DEG list
mt.genes <- grep("^MT-", RNAlistRDS005_df$gene, value=TRUE)#mitochondrial genes, no genes
rb.genes <- grep(pattern = "^RP[SL]", RNAlistRDS005_df$gene, value=TRUE)#ribosomal genes, 165 genes
RNAlistRDS005_df <- RNAlistRDS005_df %>%
  dplyr::filter(!gene %in% rb.genes) #Focus on protein coding genes

#Mutate so each row contains only a single gene in the "gene" column
gProfiler_df <- gProfiler_df %>%
  mutate(gene = strsplit(intersection, ",")) %>%
  unnest(gene) %>%
  dplyr::rename(stage = grade) %>%
  dplyr::filter(assay == "RNA")

# left_join RNAlistRDS005_df with Terms by c("cluster_id", "gene", "stage"), no multiple match warning
result_df <- left_join(RNAlistRDS005_df, gProfiler_df, by = c("cluster_id", "gene","stage"))

###Plot DEGs
# Create vulcano plot
ggplot(result_df, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(data = result_df, aes(color = "Non-overlapping"), size = 1, color = "grey") +
  geom_point(data = result_df %>% dplyr::filter(!is.na(DEG_category), DEG_category != "EndometriumTerms"),
             aes(color = DEG_category), size = 2) +
  theme_minimal() +
  facet_wrap(~stage)
ggsave(filename = paste0(dir_out, 'vulcano_DEG_category_InterestingTerms.pdf'), width =8,height=6)

# Create vulcano plot with gene labels
#logFC vs AveExpr with gene labels
InterestingTerms <- result_df[,c(1:12,34)] %>% dplyr::filter(!is.na(DEG_category), DEG_category != "EndometriumTerms") %>% 
  distinct(gene, logFC,AveExpr,p_adj.loc, .keep_all = TRUE) %>% 
  #dplyr::filter(logFC > 0.5 | logFC < -0.5) %>% 
  dplyr::filter(stage == "all", annotation_reference == "Tan", AveExpr > 18.5, logFC > 0.2)
plot <- ggplot(result_df, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(data = result_df, aes(color = "Non-overlapping"), size = 1, color = "grey") +
  geom_point(data = result_df %>% dplyr::filter(!is.na(DEG_category), DEG_category != "EndometriumTerms"),
             aes(color = DEG_category), size = 2) +
  theme_minimal()+
  ylim(0,18) +
  xlim(-.9,1.2) +
  geom_text_repel(data = InterestingTerms,
                  aes(label = gene),
                  size = 2,
                  max.overlaps = 50,
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50')
plot
ggsave(plot, filename = paste0(dir_out, 'vulcano_DEG_category_InterestingTerms_labelled.pdf'), width =16,height=12)

# Create vulcano plot with gene labels
#logFC vs AveExpr with gene labels
InterestingTerms <- result_df[,c(1:12,34)] %>% dplyr::filter(!is.na(DEG_category), DEG_category != "EndometriumTerms") %>% 
  distinct(gene, logFC,AveExpr,p_adj.loc, .keep_all = TRUE) %>% 
  #dplyr::filter(logFC > 0.5 | logFC < -0.5) %>% 
  dplyr::filter(stage == "all", annotation_reference == "Tan", abs(logFC) > 0.5)
result_df_all <- filter(result_df,stage == "all")
ggplot(result_df_all, aes(x = logFC, y = -log10(p_adj.loc))) +
  geom_point(aes(color = "Non-overlapping"), size = 1, color = "grey") +
  geom_point(data = result_df_all %>% dplyr::filter(!is.na(DEG_category), DEG_category != "EndometriumTerms"),
             aes(x = logFC, y = -log10(p_adj.loc), color = DEG_category), size = 2) +
  theme_minimal() +
  ylim(-2, 17) +
  xlim(-0.9, 1.2) +
  geom_text_repel(data = result_df_all[2:12] %>% dplyr::filter(abs(logFC) > 0.5) %>% distinct(),
                  aes(label = gene),
                  size = 2,
                  max.overlaps = 50,
                  box.padding = 0.35,
                  point.padding = 0.5,
                  segment.color = 'grey50') +
  facet_wrap(~stage)
ggsave(filename = paste0(dir_out, 'vulcano_DEG_category_InterestingTerms_labelledlogFC05.pdf'), width =8,height=6)

results_df_logFC05 <- result_df[2:12] %>% dplyr::filter(abs(logFC) > 0.5) %>% distinct()
write.csv(results_df_logFC05, paste0(dir_out,'TermslogFC05.csv'))

#logFC vs AveExpr
ggplot(result_df, aes(x = log2(AveExpr-17.00183), y = logFC)) +
  geom_point(data = result_df, aes(color = "Non-overlapping"), size = 1, color = "grey") +
  geom_point(data = result_df %>% dplyr::filter(!is.na(DEG_category), DEG_category != "EndometriumTerms"),
             aes(color = DEG_category), size = 2) +
  theme_minimal() +
  geom_text(data = result_df %>% dplyr::filter(!is.na(DEG_category), DEG_category != "EndometriumTerms"), aes(label = gene))
ggsave(filename = paste0(dir_out, 'AveExprLOG2_logFC_InterestingTerms.pdf'), width =8,height=6)

#logFC vs AveExpr with gene labels
InterestingTerms <- result_df[,c(1:12,34)] %>% dplyr::filter(!is.na(DEG_category), DEG_category != "EndometriumTerms") %>% 
  distinct(gene, logFC,AveExpr,p_adj.loc, .keep_all = TRUE) %>% 
  #dplyr::filter(logFC > 0.5 | logFC < -0.5) %>% 
  dplyr::filter(stage == "all", annotation_reference == "Tan", AveExpr > 18.5)

ggplot(result_df, aes(x = log2(AveExpr-17.00183), y = logFC)) +
  geom_point(data = result_df, aes(color = "Non-overlapping"), size = 1, color = "grey") +
  geom_point(data = InterestingTerms , aes(color = DEG_category), size = 2) +
  theme_minimal() +
  geom_text_repel(data = InterestingTerms,
                  aes(label = gene),
                  size = 2,
                  max.overlaps = 35,
                  box.padding   = 0.35, 
                  point.padding = 0.5,
                  segment.color = 'grey50')
ggsave(filename = paste0(dir_out, 'AveExprLOG2_logFC_InterestingTerms_labelled.pdf'), width =14,height=8)

#logFC vs AveExpr
ggplot(result_df, aes(x = AveExpr, y = logFC)) +
  geom_point(data = result_df, aes(color = "Non-overlapping"), size = 1, color = "grey") +
  geom_point(data = result_df %>% dplyr::filter(!is.na(DEG_category), DEG_category != "EndometriumTerms"),
             aes(color = DEG_category), size = 2) +
  theme_minimal()
 ggsave(filename = paste0(dir_out, 'AveExpr_logFC_InterestingTerms.pdf'), width =8,height=6)

>>>>

```{r setup}
library(gprofiler2)
library(plyr)
library(dplyr)
library(ggplot2)
library(tibble)
library(cowplot)
library(ggrepel)

@@

@@

gProfiler based on RNA assay and per annotation with gene expression >18.53202
```{r}
##function to run gProfiler and save output
gProfiler_FUN <- function(x,ANALYSIS,GRADE,ORDERED_QUERY,ASSAY,dir_out) {
#gProfiler
custom_bg <- unique(x$gene)
query <- x %>%
  dplyr::filter(AveExpr > 18.53202) %>%
  dplyr::filter(p_adj.loc < 0.05) %>%
  dplyr::arrange(p_adj.loc) #arrange by lowest p_adj.loc
if(nrow(query) > 0){
query <- distinct(query, gene, .keep_all= TRUE) #remove duplicates of gene with higher p_adj.loc
query <- query$gene
gostres <<- gost(query = query, 
                organism = "hsapiens", 
                ordered_query = ORDERED_QUERY, #can use TRUE, p_adj.loc has been arranged 
                multi_query = FALSE, 
                significant = TRUE, #only statistically significant results should be returned
                exclude_iea = FALSE, 
                measure_underrepresentation = FALSE, 
                evcodes = TRUE,
                user_threshold = 0.05, 
                correction_method = "g_SCS", #g_SCS is default
                domain_scope = "annotated", #if custom_bg is set, I believe this shouldn't matter too much
                custom_bg = custom_bg, #custom background, use!
                numeric_ns = "", 
                sources = NULL, #by default all sources are analyzed
                as_short_link = FALSE)
if(length(gostres) > 0){
results <- gostres$result
results$cluster_id <- levels(as.factor(x$cluster_id))
results$analysis <- print(ANALYSIS)
results$assay <- print(ASSAY)
results$grade <- print(GRADE) 
results$ordered_query <- print(ORDERED_QUERY)
results$parents <- NULL
write.csv(results,paste0("gProfiler_",ANALYSIS,"_",GRADE,"_",levels(as.factor(x$cluster_id)),"_OrderedQuery",as.character(ORDERED_QUERY),".csv"))
print(paste0("Saved gotres as gProfiler_",ANALYSIS,"_",GRADE,"_",levels(as.factor(x$cluster_id)),"_OrderedQuery",as.character(ORDERED_QUERY),".csv"))
plot <- gostplot(gostres, capped = TRUE, interactive = FALSE) + ggtitle(paste0("gProfiler_",ANALYSIS,"_",GRADE,"_",levels(as.factor(x$cluster_id)),"_OrderedQuery",as.character(ORDERED_QUERY)))
ggsave(plot = plot, filename = paste0(dir_out,"gProfiler_",ANALYSIS,"_",GRADE,"_",levels(as.factor(x$cluster_id)),"_OrderedQuery",as.character(ORDERED_QUERY),".png"), width = 7, height = 5)
print("Saved plot")
}
}
}
##function to run gProfiler_FUN on different grades, analysis and assay
RUN_gProfier <- function(GRADE, ANALYSIS,ASSAY,SPECIFICATION){
dir_out <- paste0("/home/common/data/output/projects/ENDO/E044/A031/gProfiler/E044",ANALYSIS,"/",ASSAY,"assay_",SPECIFICATION,"/")
dir.create(dir_out, recursive = TRUE)
setwd(dir_out)
#Read paths to DEG output .rds files
PATHS <- list.files(path = paste0("/home/common/data/output/projects/ENDO/E044/",ANALYSIS), pattern = "res_mm_dream.rds", recursive = TRUE, full.names = TRUE)

#Select paths based on all/mild/severe
if (GRADE == "all") { 
    PATHS <- PATHS[!grepl("_severe", PATHS)]
    PATHS <- PATHS[!grepl("_mild", PATHS)]
    print(PATHS)
    } else if (GRADE == "mild") {
    PATHS <- PATHS[grepl("_mild", PATHS)]
    print(PATHS)
    } else if  (GRADE == "severe") {
    PATHS <- PATHS[grepl("_severe", PATHS)]
    print(PATHS)
    } 
RNAlistRDS <- lapply(PATHS, readRDS)
RNAlistRDS <- unlist(RNAlistRDS, recursive = FALSE) #combine sub-lists
RNAlistRDS <- RNAlistRDS[!duplicated(RNAlistRDS)]
lapply(RNAlistRDS, gProfiler_FUN, ANALYSIS = ANALYSIS,GRADE = GRADE,ORDERED_QUERY = FALSE, ASSAY = ASSAY, dir_out = dir_out)
lapply(RNAlistRDS, gProfiler_FUN, ANALYSIS = ANALYSIS,GRADE = GRADE,ORDERED_QUERY = TRUE, ASSAY = ASSAY, dir_out = dir_out)
}

RUN_gProfier(GRADE = "all", ANALYSIS = "A030", ASSAY = "RNA",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
RUN_gProfier(GRADE = "severe", ANALYSIS = "A030", ASSAY = "RNA",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
RUN_gProfier(GRADE = "mild", ANALYSIS = "A030", ASSAY = "RNA",SPECIFICATION = "PerAnnotation_AveExpr18_53202")
```

@@

@@
